Component response rate variation underlies the stability of complex systems

### 2020-03-21 notebook/ms.Rmd
Rearrange the results files so that the manuscript calls them from the inst Rpackage folder

<!---
Abstract current word count: 225
Text current word count: 1453
--->

@@

<!---

NOTE: Consider looking at the resilience of the stable networks given $\gamma = 1$ and $Var(\gamma)$, which is defined by the inverse of the leading eigenvalue. The inverse of the real part of the leading eigenvalue indicates how quickly the system returns to equilibrium (resilience). See @Landi2018 for more. Intuitively, resilience should perhaps increase a bit if we're only looking at the subset of systems that are stable (many of which will have had their eigenvalues shifted left)?

Landi, P., Minoarivelo, H. O., Brännström, Å., Hui, C., & Dieckmann, U. (2018). Complexity and stability of ecological networks: A review of the theory. Population Ecology, 0(0), In press. https://doi.org/10.1007/s10144-018-0628-3

--->

<!---

1. As above, check out the change in resilience defined as the difference beween the inverse of the real part of the leading eigenvalue in $Var(\gamma)$ versus $\gamma = 1$. Initial simulaitons suggest that resilience is increased on average by $Var(\gamma)$, and increasingly so as S increases.

2. Check for induced correlations causing stabilisation (and E1, E2, and EC, more generally), as per Reviewer 2 specific comment 3. If possible, give an analytical explanation for the increase in stability that has to do with the error inducing changes in correlations.

--->

@@

```{r echo = FALSE}
S     <- 200;
C     <- 0.05;
sigma <- 0.4;
pr_st <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
pr_st <- pr_st[,-1];
```

```{r, echo = FALSE}
A0 <- read.csv(file = "sim_results/bi_gamma/S200_A0.csv");
A0 <- as.matrix(A0[,-1]);
A1 <- read.csv(file = "sim_results/bi_gamma/S200_A1.csv");
A1 <- as.matrix(A1[,-1]);

A0_e   <- eigen(A0)$values;
A0_r   <- Re(A0_e);
A0_i   <- Im(A0_e);
A1_e   <- eigen(A1)$values;
A1_r   <- Re(A1_e);
A1_i   <- Im(A1_e);

A0_vm       <- A0;
diag(A0_vm) <- NA;
A0vec       <- as.vector(t(A0_vm));
A0vec       <- A0vec[is.na(A0vec) == FALSE];
A1_vm       <- A1;
diag(A1_vm) <- NA;
A1vec       <- as.vector(t(A1_vm));
A1vec       <- A1vec[is.na(A1vec) == FALSE];
fhalf       <- 1:(0.5*length(A1vec));
shalf       <- (0.5*length(A1vec)+1):length(A1vec);
pr_st       <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
pr_st       <- pr_st[,-1];
```

@@

```{r, echo = FALSE}
dat <- read.csv(file = "sim_results/C_1/random_all.csv");
dat <- dat[,-1];
```

@@

<!---

For all simulated network structures, interaction strengths between components were sampled as in random networks ($\mathcal{N}(0, 0.4^{2})$), but algorithms for building networks necessarily resulted in variable $C$.

--->

@@

<!--- Tempting to remove this section entirely, but if there is enough space for it, then maybe retain it; reviewers did not mention a problem with it, and it's a short and relevant point to make.  In any case, it should be swapped with the section on feasibility below. --->
**Targeted manipulation of $\boldsymbol{\gamma}$**. To further investigate the potential of $\sigma^{2}_{\gamma}$ to be stabilising, I used a genetic algorithm. Genetic algorithms are heuristic tools that mimic evolution by natural selection, and are useful when the space of potential solutions (in this case, possible combinations of $\gamma$ values leading to stability in a complex system) is too large to search exhaustively [@Hamblin2013]. Generations of selection on $\gamma$ value combinations to minimise $\Re(\lambda_{max})$ demonstrated the potential for $\sigma^{2}_{\gamma}$ to increase system stability. Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were up to four orders of magnitude higher than when $\gamma = 1$ (see Supplementary Information), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $\mathbf{M}$ elements (i.e., by manipulating component response rates rather than interactions between components).

@@

<!---  (in practice, diagonal elements of $\textbf{M}$ were standardised so that mean values were identical before and after adding $\gamma$) --->

@@

```{r, echo = FALSE, fig.height = 6, fig.width = 6}
sim_rhos <- read.csv(file = "sim_results/rhos/sim_rhos.csv");
par(mar = c(5, 5, 1, 1));
plot(x = sim_rhos[,1], y = sim_rhos[,20], type = "b", pch = 20, lwd = 2,
     xlab = expression(paste("Correlation between A"[ij]," and A"[ji], 
                             " (",rho,")")),
     ylab = "Real part of leading eigenvalue", col = "dodgerblue4",
     ylim = c(min(sim_rhos[,22:25]), max(sim_rhos[,22:25])),
     cex.lab = 1.25, cex.axis = 1.25);
mbox <- function(x0, x1, y0, y1){
    xx <- seq(from=x0, to=x1, length.out = 100);
    yy <- seq(from=y0, to=y1, length.out = 100);
    xd <- c(rep(x0, 100), xx, rep(x1,100), rev(xx));
    yd <- c(yy, rep(y1,100), rev(yy), rep(y0, 100));
    return(list(x=xd, y=yd));
}
more_stable_1 <- sim_rhos[,21] - sim_rhos[,20]
more_stable_2 <- which(more_stable_1 < 0);
min_x         <- min(more_stable_2);
max_x         <- max(more_stable_2);
shade_box     <- mbox(x0 = sim_rhos[min_x, 1], x1 = sim_rhos[max_x, 1], 
                      y0 = -12, y1 = 15);
polygon(x = shade_box$x, y = shade_box$y, col = "grey80", border = NA);
abline(h = 0, lwd = 0.8, lty = "dotted", col = "black");
for(i in 1:dim(sim_rhos)[1]){
    lines(x = c(sim_rhos[i,1], sim_rhos[i,15]), col = "purple",
          y = c(sim_rhos[i,20], sim_rhos[i,21]), lwd = 1.5);
}
arrows(x0 = sim_rhos[,1], x1 = sim_rhos[,1], y0 = sim_rhos[,20], 
       y1 = sim_rhos[,22], angle = 90, length = 0.05, lwd = 1, col = "dodgerblue4");
arrows(x0 = sim_rhos[,1], x1 = sim_rhos[,1], y0 = sim_rhos[,20], 
       y1 = sim_rhos[,23], angle = 90, length = 0.05, lwd = 1, col = "dodgerblue4");
arrows(x0 = sim_rhos[,15], x1 = sim_rhos[,15], y0 = sim_rhos[,21], 
       y1 = sim_rhos[,24], angle = 90, length = 0.05, lwd = 1, col = "firebrick");
arrows(x0 = sim_rhos[,15], x1 = sim_rhos[,15], y0 = sim_rhos[,21], 
       y1 = sim_rhos[,25], angle = 90, length = 0.05, lwd = 1, col = "firebrick");
points(x = sim_rhos[,1], y = sim_rhos[,20], type = "b", pch = 20, lwd = 3,
       col = "dodgerblue4");
points(x = sim_rhos[,15], y = sim_rhos[,21], type = "b", pch = 15, lwd = 3,
       col = "firebrick");
points(x = sim_rhos[,1], y = sim_rhos[,20], type = "l", pch = 20, lwd = 3,
       col = "dodgerblue4");
points(x = sim_rhos[,15], y = sim_rhos[,21], type = "l", pch = 15, lwd = 3,
       col = "firebrick");
legend("topleft", fill = c("dodgerblue4", "firebrick"), cex = 1.55,
       legend = c(expression(paste(sigma[gamma]^2," = 0")), 
                  expression(paste(sigma[gamma]^2," = 1/3"))),
       bg = "white", y.intersp = 1.5);
box();
# Note that the 1/3 comes from the definition of unif distr. from 0 to 2.
```

@@

```{r, echo = FALSE, fig.height = 8, fig.width = 6}
sim10 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim10.csv");
sim15 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim15.csv");
sim20 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim20.csv");
sim25 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim25.csv");
sim30 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim30.csv");
sim35 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim35.csv");
sim40 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim40.csv");
sim45 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim45.csv");

@@

```{r, echo = FALSE}
dat <- read.csv(file = "sim_results/C_1/random_all.csv");
dat <- dat[,-1];
```

@@

```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
dat <- read.csv(file = "sim_results/C_1/random_all.csv");
dat <- dat[,-1];
plot_stables <- function(dat, S_s = 32){
    Ns          <- 1:S_s;
    par(oma = c(6, 6, 1, 6), mar = c(0.5, 0.5, 0.5, 0.5));
    #=================================
    bar_dat                      <- t(cbind(dat[Ns,3], dat[Ns,5]));
    log_bar_dat                  <- log(bar_dat);
    log_bar_dat[log_bar_dat < 0] <- 0; 
    barplot(log_bar_dat, beside = TRUE, col = c("dodgerblue4", "firebrick"),
            names.arg = dat[Ns,1], ylim = c(0, 16), xlab = "",
            ylab = "Ln number of stable communities", cex.lab = 1, 
            cex.axis = 1.25, xlim = c(1, 94), cex.names = 1, yaxt = "n");
    axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12, 14), cex.axis = 1.5);
    box(lwd = 2);
    par(new = TRUE);
    y1     <- dat[1:S_s,6] / (dat[1:S_s,5]);
    x1     <- seq(from = 2.132, to = 15.1112, length = S_s);
    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1.1),
         xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, 
         cex = 1, col = "black", yaxs="i");
    points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
    axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
    legend("topleft", c(expression(paste(sigma[gamma]^2," = 0")), 
                        expression(paste(sigma[gamma]^2," = 1/3"))), 
           pch=15, col=c("dodgerblue4","firebrick"), cex = 1.5, horiz = TRUE);
    #=================================
    mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 
          line = 3.0);
    mtext(side = 2, text = "Ln number of stable systems", cex = 2, 
          outer = TRUE, line = 3.5);
    mtext(side = 4, text = expression(
        paste("Pr. of systems stable due to Var(",gamma,")")),
        cex = 2, outer = TRUE, line = 3.5);
}
plot_stables(dat);
```

>>>>

<!---
Abstract current word count: 
Text current word count: 
--->

@@

@@

```{r echo = FALSE}
S     <- 200;
C     <- 0.05;
sigma <- 0.4;
```

@@

```{r, echo = FALSE}
dat <- read.csv(file = "../inst/extdata/random_all.csv");
```

@@

@@

**Targeted manipulation of $\boldsymbol{\gamma}$**. To further investigate the potential of $\sigma^{2}_{\gamma}$ to be stabilising, I used a genetic algorithm. Genetic algorithms are heuristic tools that mimic evolution by natural selection, and are useful when the space of potential solutions (in this case, possible combinations of $\gamma$ values leading to stability in a complex system) is too large to search exhaustively [@Hamblin2013]. Generations of selection on $\gamma$ value combinations to minimise $\Re(\lambda_{max})$ demonstrated the potential for $\sigma^{2}_{\gamma}$ to increase system stability. Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were up to four orders of magnitude higher than when $\gamma = 1$ (see Supplementary Information), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $\mathbf{M}$ elements (i.e., by manipulating component response rates rather than interactions between components).

@@

@@

```{r, echo = FALSE, fig.height = 6, fig.width = 6}
sim_rhos <- read.csv(file = "../inst/extdata/rhos/sim_rhos.csv");
par(mar = c(5, 5, 1, 1));
plot(x = sim_rhos[,1], y = sim_rhos[,20], type = "b", pch = 20, lwd = 2,
     xlab = expression(paste("Correlation between A"[ij]," and A"[ji], 
                             " (",rho,")")),
     ylab = "Real part of leading eigenvalue", col = "dodgerblue4",
     ylim = c(min(sim_rhos[,22:25]), max(sim_rhos[,22:25])),
     cex.lab = 1.25, cex.axis = 1.25);
mbox <- function(x0, x1, y0, y1){
    xx <- seq(from=x0, to=x1, length.out = 100);
    yy <- seq(from=y0, to=y1, length.out = 100);
    xd <- c(rep(x0, 100), xx, rep(x1,100), rev(xx));
    yd <- c(yy, rep(y1,100), rev(yy), rep(y0, 100));
    return(list(x=xd, y=yd));
}
more_stable_1 <- sim_rhos[,21] - sim_rhos[,20]
more_stable_2 <- which(more_stable_1 < 0);
min_x         <- min(more_stable_2);
max_x         <- max(more_stable_2);
shade_box     <- mbox(x0 = sim_rhos[min_x, 1], x1 = sim_rhos[max_x, 1], 
                      y0 = -12, y1 = 15);
polygon(x = shade_box$x, y = shade_box$y, col = "grey80", border = NA);
abline(h = 0, lwd = 0.8, lty = "dotted", col = "black");
for(i in 1:dim(sim_rhos)[1]){
    lines(x = c(sim_rhos[i,1], sim_rhos[i,15]), col = "purple",
          y = c(sim_rhos[i,20], sim_rhos[i,21]), lwd = 1.5);
}
arrows(x0 = sim_rhos[,1], x1 = sim_rhos[,1], y0 = sim_rhos[,20], 
       y1 = sim_rhos[,22], angle = 90, length = 0.05, lwd = 1, col = "dodgerblue4");
arrows(x0 = sim_rhos[,1], x1 = sim_rhos[,1], y0 = sim_rhos[,20], 
       y1 = sim_rhos[,23], angle = 90, length = 0.05, lwd = 1, col = "dodgerblue4");
arrows(x0 = sim_rhos[,15], x1 = sim_rhos[,15], y0 = sim_rhos[,21], 
       y1 = sim_rhos[,24], angle = 90, length = 0.05, lwd = 1, col = "firebrick");
arrows(x0 = sim_rhos[,15], x1 = sim_rhos[,15], y0 = sim_rhos[,21], 
       y1 = sim_rhos[,25], angle = 90, length = 0.05, lwd = 1, col = "firebrick");
points(x = sim_rhos[,1], y = sim_rhos[,20], type = "b", pch = 20, lwd = 3,
       col = "dodgerblue4");
points(x = sim_rhos[,15], y = sim_rhos[,21], type = "b", pch = 15, lwd = 3,
       col = "firebrick");
points(x = sim_rhos[,1], y = sim_rhos[,20], type = "l", pch = 20, lwd = 3,
       col = "dodgerblue4");
points(x = sim_rhos[,15], y = sim_rhos[,21], type = "l", pch = 15, lwd = 3,
       col = "firebrick");
legend("topleft", fill = c("dodgerblue4", "firebrick"), cex = 1.55,
       legend = c(expression(paste(sigma[gamma]^2," = 0")), 
                  expression(paste(sigma[gamma]^2," = 1/3"))),
       bg = "white", y.intersp = 1.5);
box();
# Note that the 1/3 comes from the definition of unif distr. from 0 to 2.
```

@@

```{r, echo = FALSE, fig.height = 7.5, fig.width = 6}
sim10 <- read.csv(file = "../inst/extdata/rhos/sim10.csv");
sim15 <- read.csv(file = "../inst/extdata/rhos/sim15.csv");
sim20 <- read.csv(file = "../inst/extdata/rhos/sim20.csv");
sim25 <- read.csv(file = "../inst/extdata/rhos/sim25.csv");
sim30 <- read.csv(file = "../inst/extdata/rhos/sim30.csv");
sim35 <- read.csv(file = "../inst/extdata/rhos/sim35.csv");
sim40 <- read.csv(file = "../inst/extdata/rhos/sim40.csv");
sim45 <- read.csv(file = "../inst/extdata/rhos/sim45.csv");

@@

@@

```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
plot_stables <- function(dat, S_s = 32){
    Ns          <- 1:S_s;
    par(oma = c(6, 6, 1, 6), mar = c(0.5, 0.5, 0.5, 0.5));
    #=================================
    bar_dat                      <- t(cbind(dat[Ns,3], dat[Ns,5]));
    log_bar_dat                  <- log(bar_dat);
    log_bar_dat[log_bar_dat < 0] <- 0; 
    barplot(log_bar_dat, beside = TRUE, col = c("dodgerblue4", "firebrick"),
            names.arg = dat[Ns,1], ylim = c(0, 16), xlab = "",
            ylab = "Ln number of stable communities", cex.lab = 1, 
            cex.axis = 1.25, xlim = c(1, 94), cex.names = 1, yaxt = "n");
    axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12, 14), cex.axis = 1.5);
    box(lwd = 2);
    par(new = TRUE);
    y1     <- dat[1:S_s,6] / (dat[1:S_s,5]);
    x1     <- seq(from = 2.132, to = 15.1112, length = S_s);
    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1.1),
         xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, 
         cex = 1, col = "black", yaxs="i");
    points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
    axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
    legend("topleft", c(expression(paste(sigma[gamma]^2," = 0")), 
                        expression(paste(sigma[gamma]^2," = 1/3"))), 
           pch=15, col=c("dodgerblue4","firebrick"), cex = 1.5, horiz = TRUE);
    #=================================
    mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 
          line = 3.0);
    mtext(side = 2, text = "Ln number of stable systems", cex = 2, 
          outer = TRUE, line = 3.5);
    mtext(side = 4, text = expression(
        paste("Pr. of systems stable due to Var(",gamma,")")),
        cex = 2, outer = TRUE, line = 3.5);
}
plot_stables(dat);
```
